### 2016-10-13 Paper_v4.tex
Added the AIC blurb

\subsection*{Estimating an ensemble of reduced order complement models.}
A critical challenge for the development of any dynamic model is the estimation of model parameters.
We estimated an ensemble of complement model parameters using \textit{in~vitro} time-series data sets generated with and without zymosan, a lectin pathway activator \cite{morad2015time}.
The residual between model simulations and experimental measurements was minimized using the Pareto Optimal Ensemble Technique (JuPOETs) \cite{JuPOETs-BioArXiv}
starting from a initial guess generated by the dynamic optimization with particle swarms (DOPS) routine.
Unless otherwise specified, all initial conditions were assumed to be at their mean physiological values.
While we had significant training data, the parameter estimation problem was underdetermined (we were not able to uniquely determine model parameters).
Thus, instead of using the best-fit yet uncertain parameter set, we estimated an ensemble of probable parameter sets to quantify model uncertainty
(N = 2100, see materials and methods).
The complement model ensemble captured the behavior of both the alternate and lectin pathways (Fig. \ref{fig-fit}).
To estimate alternate pathway model parameters, we used C3a and C5a measurements in the absence of zymosan (Fig. \ref{fig-fit}A and B).
On the other hand, lectin pathway parameters were estimated from C3a and C5a measurements in the presence of 1mg/ml zymosan (Fig. \ref{fig-fit}C and D).
The reduced order model reproduced a panel of alternate and lectin pathway data sets in the neighborhood of physiological factor and inhibitor concentrations.
However, it was unclear whether the reduced order model could predict new data, without updating the model parameters.
To address this question, we fixed the model parameters and simulated data sets not used for model training.

@@

We tested the predictive power of the reduced order complement model with data not used during model training (Fig. \ref{fig-prediction}).
Six validation cases were considered, three for C3a and C5a, respectively.
All model parameters and initial conditions were fixed for the validation simulations (with the exception of zymosan, and other experimentally mandated changes).
The ensemble of reduced order models predicted the qualitative dynamics of C3a formation (Fig. \ref{fig-prediction}, top),
and C5a formation (Fig. \ref{fig-prediction}, bottom) at three inducer concentrations.
The rate of C3a formation and C3a peak time were directly proportional to initiator dose.
Similarly, the C5a plateau and rate of formation were also directly proportional to initiator dose, with the lag time being indirectly proportional to initiator exposure for both C3a and C5a.
However, there were shortcomings with model performance.
First, while the overall C3a trend was captured (within the 99\% confidence interval), the C3a dynamics were too fast with the exception of the low dose case.
We believe the C3a time scale was related to our choice of training data, how we modeled the tickover mechanism, and factor B and D limitation.
We trained the model using either no or 1 mg/ml zymosan, but predicted cases in a different initiator range;
comparing training to prediction, the model performance e.g., the shape of the C3a trajectory was biased towards either high or very low initiator doses.
Next, tickover was modeled as a first-order generation processes where C3wBb formation and activity was lumped into the AP C3 convertase.
Thus, we skipped an important upstream step which could influence AP C3 convertase formation by attenuating the rate C3 cleavage into C3a and C3b.
We also assumed both factor B and factor D were not limiting, thereby artificially accelerating the rate of AP C3 convertase formation.
The C5a predictions followed a similar trend as C3a; we captured the long-time C5a behavior but over predicted the time scale of C5 cleavage.
However, because the C5a time scale depends strongly upon C3 convertase formation, we can likely correct the C5 issues by fixing the rate of C3 cleavage.
Despite these shortcomings, we qualitatively predicted experimental measurements not used for model training typically within the 99\% confidence of the ensemble, for three inducer levels.
Next, we used global sensitivity and robustness analysis to determine which parameters and species controlled the performance of the complement model.

@@

Pairwise parameter perturbations identified modes of crosstalk which controlled fibrinolysis model performance (Fig. \ref{fig-pairwise-analysis}).
We perturbed each pairwise parameter combination by 10\%, and calculated the distance between the perturbed and nominal state for each parameter set in the ensemble.
We then clustered the mean response of each parameter combination based upon the euclidian distance between the perturbed and nominal states into low (green), medium (red) and high (blue) response clusters.
Four of the 28 parameters (or approximately 14\% of the overall model parameters) were in the high response cluster (Fig. \ref{fig-pairwise-analysis}, blue cluster).
These parameters included the rate constant controlling the basal formation of C3b (\#12), C3a degradation (\#26) as well as the catalytic rate constant governing CP C3 convertase activity (\#22).
The only C5 related parameter in the high response group was the rate constant controlling the formation of CP C5 convertase (\#15).
Approximately, 36\%, or 10 of the 28 model parameters, were clustered in the medium impact cluster (Fig. \ref{fig-pairwise-analysis}, red cluster).
Three parameters (\#10,~\#1,~\#27) were especially important in the medium impact cluster.
The reaction order governing CP C3 convertase activity was important (\#10),
along with the catalytic rate constant controlling the formation of C4a and C4b in the lectin initiation pathway (\#1), and the constant controlling the inhibitory action of C4BP (\#27).
Lastly,  50\% of the model parameters were clustered in the low response cluster (Fig. \ref{fig-pairwise-analysis}, green cluster).
Many of these parameters influenced complement activation; for example, parameter \#23 (the K$_M$ of CP C3 convertase) was important, just not to the extent of the other parameters.
The pairwise perturbation study also identified synergistic crosstalk between parameters.
For example, in the high impact cluster, three synergistic relationships were identified, a single positive and two negative cases.
Parameters \#12 (rate constant governing basal C3b formation) and \#15 (formation of CP C5 conevertase) acted synergistically to increase the system response.
On the other hand, simultaneous changes in parameters \#12 and \#22 or \#15 and \#26 decreased the system response relative to a single perturbation.
However, sensitivity coefficients quantify how changes in parameters e.g., rate or saturation constants affect model performance.
To more closely simulate a clinical intervention e.g., administration of anti-complement inhibitors, we performed robustness analysis.

@@

In this study, we developed an ensemble of experimentally validated reduced order complement models using multiobjective optimization.
The modeling approach combined ordinary differential equations with logical rules to produce a complement model with a limited number of equations and parameters.
The reduced order model,  which described the lectin and alternative pathways, consisted of 18 differential equations with 28 parameters.
Thus, the model was an order of magnitude smaller and included more pathways than comparable mathematical models in the literature.
We estimated an ensemble of model parameters from \textit{in~vitro} time series measurements of the C3a and C5a complement proteins.
Subsequently, we validated the model on unseen C3a and C5a measurements that were not used for model training.
Despite its small size, the model was surprisingly predictive.
After validation, we performed global sensitivity and robustness analysis to estimate which parameters and species
controlled model performance. These analyses suggested complement was robust to any single therapeutic intervention.
The only intervention that consistently reduced C3a and C5a formation for all cases was a knockdown of both C3 and C5.
Taken together, we developed a reduced order complement model that was computationally inexpensive,
and could easily be incorporated into pre-existing or new pharmacokinetic models of immune system function.
The model described experimental data, and predicted the need for multiple points of intervention to disrupt complement activation.

>>>>

\subsection*{Estimating an ensemble of reduced order complement models.}
A critical challenge for the development of any dynamic model is the estimation of model parameters.
We estimated an ensemble of complement model parameters using \textit{in~vitro} time-series data sets generated with and without zymosan, a lectin pathway activator \cite{morad2015time}.
The residual between model simulations and experimental measurements was minimized using the Pareto Optimal Ensemble Technique (JuPOETs) \cite{JuPOETs-BioArXiv}
starting from a initial guess generated by the dynamic optimization with particle swarms (DOPS) routine.
Unless otherwise specified, all initial conditions were assumed to be at their mean physiological values.
While we had significant training data, the parameter estimation problem was underdetermined (we were not able to uniquely determine model parameters).
Thus, instead of using the best-fit yet uncertain parameter set, we estimated an ensemble of probable parameter sets to quantify model uncertainty
(N = 2100, see materials and methods).
The complement model ensemble captured the behavior of both the alternate and lectin pathways (Fig. \ref{fig-fit}).
To estimate alternate pathway model parameters, we used C3a and C5a measurements in the absence of zymosan (Fig. \ref{fig-fit}A and B).
On the other hand, lectin pathway parameters were estimated from C3a and C5a measurements in the presence of 1mg/ml zymosan (Fig. \ref{fig-fit}C and D).
The reduced order model reproduced a panel of alternate and lectin pathway data sets in the neighborhood of physiological factor and inhibitor concentrations.
The model fit for parameter sets estimated by JuPOETs, quantified by the Akaike information criterion (AIC), was statistically significantly different than a random parameter control
for each case at a 95\% confidence level. However, it was unclear whether the reduced order model could predict new data, without updating the model parameters.
To address this question, we fixed the model parameters and simulated data sets not used for model training.

@@

We tested the predictive power of the reduced order complement model with data not used during model training (Fig. \ref{fig-prediction}).
Six validation cases were considered, three for C3a and C5a each, respectively.
Similar to model training, we compared the AIC for each prediction case to a randomized parameter family.
All model parameters and initial conditions were fixed for the validation simulations (with the exception of zymosan, and other experimentally mandated changes).
The ensemble of reduced order models predicted the qualitative dynamics of C3a formation (Fig. \ref{fig-prediction}, top),
and C5a formation (Fig. \ref{fig-prediction}, bottom) at three inducer concentrations.
For each training case, the AIC was statistically significantly different than the random parameter control for a 95\% confidence level.
The rate of C3a formation and C3a peak time were directly proportional to initiator dose.
Similarly, the C5a plateau and rate of formation were also directly proportional to initiator dose, with the lag time being indirectly proportional to initiator exposure for both C3a and C5a.
However, there were shortcomings with model performance.
First, while the overall C3a trend was captured (within the 99\% confidence interval), the C3a dynamics were too fast with the exception of the low dose case.
We believe the C3a time scale was related to our choice of training data, how we modeled the tickover mechanism, and factor B and D limitation.
We trained the model using either no or 1 mg/ml zymosan, but predicted cases in a different initiator range;
comparing training to prediction, the model performance e.g., the shape of the C3a trajectory was biased towards either high or very low initiator doses.
Next, tickover was modeled as a first-order generation processes where C3wBb formation and activity was lumped into the AP C3 convertase.
Thus, we skipped an important upstream step which could influence AP C3 convertase formation by attenuating the rate C3 cleavage into C3a and C3b.
We also assumed both factor B and factor D were not limiting, thereby artificially accelerating the rate of AP C3 convertase formation.
The C5a predictions followed a similar trend as C3a; we captured the long-time C5a behavior but over predicted the time scale of C5 cleavage.
However, because the C5a time scale depends strongly upon C3 convertase formation, we can likely correct the C5 issues by fixing the rate of C3 cleavage.
Despite these shortcomings, we qualitatively predicted experimental measurements not used for model training typically within the 99\% confidence of the ensemble, for three inducer levels.
Next, we used global sensitivity and robustness analysis to determine which parameters and species controlled the performance of the complement model.

@@

Pairwise parameter perturbations identified crosstalk within the complement model (Fig. \ref{fig-pairwise-analysis}).
We perturbed each pairwise combination of parameters by 10\%, and calculated the distance between the perturbed and nominal state for each parameter set in the ensemble.
We then clustered the mean response of each parameter combination based upon the euclidian distance between the perturbed and nominal states into low (green),
medium (red) and high (blue) response clusters. A low response (white) meant the parameter perturbations did not significantly change the system state compared with the nominal case.
Four of the 28 parameters (or approximately 14\% of the overall model parameters) were in the high response cluster (Fig. \ref{fig-pairwise-analysis}, blue cluster).
These parameters included the rate constant controlling the basal formation of C3b (\#12), C3a degradation (\#26) as well as the catalytic rate constant governing CP C3 convertase activity (\#22).
The only C5 related parameter in the high response group was the rate constant controlling the formation of CP C5 convertase (\#15).
Approximately, 36\%, or 10 of the 28 model parameters, were clustered in the medium impact cluster (Fig. \ref{fig-pairwise-analysis}, red cluster).
Three parameters (\#10,~\#1,~\#27) were especially important in this cluster;
The reaction order governing CP C3 convertase activity was important (\#10),
along with the rate constant controlling C4a and C4b formation from C4 in the lectin initiation pathway (\#1), and the constant controlling the inhibitory action of C4BP (\#27).
Lastly,  50\% of the model parameters were clustered in the low response cluster (Fig. \ref{fig-pairwise-analysis}, green cluster).
Many of these parameters influenced complement activation; for example, parameter \#23 (the CP C3 convertase saturation constant) was important, just not to the extent of other model parameters.
Pairwise synergistic interactions between parameters were also identified.
For example, in the high impact cluster, three synergistic relationships were identified, a single positive and two negative cases.
Parameters \#12 (rate constant governing basal C3b formation) and \#15 (formation of CP C5 conevertase) acted synergistically to increase the system response.
On the other hand, simultaneously changing parameters \#12 and \#22 or \#15 and \#26 decreased the system response relative to a single perturbation.
However, the most striking examples of synergy occurred in the medium impact cluster; for example, simultaneously increasing parameters \#13 (rate constant governing AP C3 convertase formation)
and \#19 (saturation constant governing AP C5 convertase activity) significantly changed the model state.
Changes in parameter \#3 (rate constant governing C2a and C2b formation from C2) showed both positive and negative synergistic effects
depending upon the other parameter that was perturbed.
Taken together, [FINISH ME].
However, sensitivity coefficients quantify how changes in parameters e.g., rate or saturation constants affect model performance.
To more closely simulate a clinical intervention e.g., administration of anti-complement inhibitors, we performed robustness analysis in the absence and presence of flow.

@@

In this study, we estimated an ensemble of experimentally validated reduced order complement models using multiobjective optimization.
The modeling approach combined ordinary differential equations with logical rules to produce a complement model with a limited number of equations and parameters.
The reduced order model,  which described the lectin and alternative pathways, consisted of 18 differential equations with 28 parameters.
Thus, the model was an order of magnitude smaller and included more pathways than comparable mathematical models in the literature.
We estimated an ensemble of model parameters from \textit{in~vitro} time series measurements of the C3a and C5a complement proteins.
Subsequently, we validated the model on unseen C3a and C5a measurements that were not used for model training.
Despite its small size, the model was surprisingly predictive.
After validation, we performed global sensitivity and robustness analysis to estimate which parameters and species
controlled model performance. These analyses suggested complement was robust to any single therapeutic intervention.
The only intervention that consistently reduced C3a and C5a formation for all cases was a knockdown of both C3 and C5.
Taken together, we developed a reduced order complement model that was computationally inexpensive,
and could easily be incorporated into pre-existing or new pharmacokinetic models of immune system function.
The model described experimental data, and predicted the need for multiple points of intervention to disrupt complement activation.
